---
title: "concatstring"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{concatstring}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(concatstring)
```

## Introduction

The `concatstring` package is a lightweight R package designed to combine two character vectors elementwise using `paste()`, returning the result as a structured object of class `concat_result`.

### Why use this package?

This package provides:

* A custom S3 class (`concat_result`) to wrap the result
* A `print()` method for clean console output
* A `summary()` method for descriptive summaries
* A `plot()` method for visualising string lengths

It’s intended for teaching, lightweight text analysis, or wrapping basic string operations in a reusable and extensible format.

---

## Basic usage

You can combine two strings using:

```{r basic}
concat_strings("Hello", "World")
```

This returns an object of class `concat_result`, which contains:

* The original inputs `x` and `y`
* The concatenated result `result`
* The class `"concat_result"`

---

## Vector input

The function is vectorised, meaning it works elementwise over character vectors of the same length:

```{r vector}
concat_strings(c("Fast", "Strong"), c("Car", "Wind"))
```

---

## Output structure

You can inspect the structure of a `concat_result` object using `str()`:

```{r structure}
res <- concat_strings("Hello", "World")
str(res)
```

---

## S3 methods

The package includes print, summary, and plot methods specific to `concat_result`.

### Print

```{r print}
print(res)
```

### Summary

```{r summary}
summary(res)
```

This tells you how many elements were concatenated and shows the first few results.

---

### Plot

When the result contains multiple elements, the `plot()` method creates a barplot of string lengths:

```{r plot}
multi <- concat_strings(c("Quick", "Tiny", "Solid"), c("Fox", "Stone", "Wall"))
plot(multi)
```

If only one concatenated string is present, a warning is shown:

```{r plot-warning}
single <- concat_strings("One", "Word")
plot(single)
```

---

## Conclusion

This package demonstrates how basic string operations can be extended using S3 object-oriented programming in R. It’s ideal for learning how to wrap functionality into reusable and testable code, and for showing how to build a minimal, clean R package with proper testing, documentation, and extensibility.

---

## License

MIT © 2025

```